Miniproject - Algebraic Geometry

### 2021-12-07 Algebraic geometry/submission.tex
Wrote sage script for alg geo

\newtheorem{definition}{Definition}
\newtheorem{lemma}[definition]{Lemma}
\newtheorem{example}[definition]{Example}
\newtheorem{proposition}[definition]{Proposition}
\newtheorem{corollary}[definition]{Corollary}

@@

    It follows that $\dim(\extpow^d k^n) = \dim(\Proj^{\{1, ..., n\}^{(d)}})$ and we find that $f$ is also injective.
\end{proof}
\begin{corollary}[2b]
    Let $\bar{f}: \Proj(\extpow^d k^n) \to \Proj^{{n \choose d} - 1}$ be the map $f$ from before modulo $k^*$.
    Then
    \begin{equation*}
        \rho = f \circ \phi
    \end{equation*}
    and in particular, we see that $\phi(Gr(d, n))$ is a projective variety and isomorphic to $\rho(\Gr(d, n))$.
\end{corollary}
\begin{proposition}[2c]
    The map $\phi$ is injective.
\end{proposition}
\begin{proof}
    Consider two $d$-dimensional subspaces $U, W$ of $k^n$ with $\phi(U) = \phi(W)$.
    Let $u_1, ..., u_l$ be a basis of $U \cap W$ and extend it to bases $u_1, ..., u_d$ of $U$ and $u_1, ..., u_l, w_{l + 1}, ..., w_d$ of $W$.
    As $\phi(U) = \phi(W)$, we can assume that the $u_i, w_i$ are scaled such that
    \begin{align*}
        0 =& (u_1 \wedge ... \wedge u_d) - (u_1 \wedge ... \wedge u_l \wedge w_{l + 1} \wedge ... \wedge w_d) \\
        =& u_1 \wedge ... \wedge u_l \wedge \bigl( (u_{l + 1} \wedge ... \wedge u_d) - (w_{l + 1} \wedge ... \wedge w_d) \bigr)
    \end{align*}
    By Lemma~\ref{prop:subspace_closed} we see that
    \begin{equation*}
        u_2 \wedge ... \wedge u_l \wedge \bigl( (u_{l + 1} \wedge ... \wedge u_d) - (w_{l + 1} \wedge ... \wedge w_d) \bigr) = 0
    \end{equation*}
    as $u_1 \notin \vspan\{u_2, ..., u_d, w_{l + 1}, ..., w_d\}$.
    Inductively, this argument shows that
    \begin{equation*}
        (u_{l + 1} \wedge ... \wedge u_d) - (w_{l + 1} \wedge ... \wedge w_d) = 0
    \end{equation*}
    If $l < d$, we can now apply Lemma~\ref{prop:subspace_closed} again to see that
    \begin{equation*}
        u_{l + 1} \in \vspan\{ u_{l + 2}, ..., u_d, w_{l + 1}, ..., w_d \}
    \end{equation*}
    as $u_{l + 2} \wedge ... \wedge u_d \neq 0$ by Lemma~\ref{prop:extpow_zero_iff_independent}.
    However, this contradicts the linear independence of $u_{l + 1}, ..., u_d, w_{l + 1}, ..., w_d$.
    Hence it must be $l = d$ and so $U = W$.
\end{proof}
\end{document}

>>>>

\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{lemma}[definition]{Lemma}
\newtheorem{example}[definition]{Example}
\newtheorem{proposition}[definition]{Proposition}
\newtheorem{corollary}[definition]{Corollary}

@@

    It follows that $\dim(\extpow^d k^n) = \dim(\Proj^{\{1, ..., n\}^{(d)}})$ and we find that $f$ is also injective.
\end{proof}
\begin{corollary}[2b]
    \label{prop:image_phi}
    Let $\bar{f}: \Proj(\extpow^d k^n) \to \Proj^{{n \choose d} - 1}$ be the map $f$ from before modulo $k^*$.
    Then
    \begin{equation*}
        \rho = f \circ \phi
    \end{equation*}
    and in particular, we see that $\phi(Gr(d, n))$ is a projective variety and isomorphic to $\rho(\Gr(d, n))$.
\end{corollary}
\begin{proposition}[2c]
    The map $\phi$ is injective.
\end{proposition}
\begin{proof}
    Consider two $d$-dimensional subspaces $U, W$ of $k^n$ with $\phi(U) = \phi(W)$.
    Let $u_1, ..., u_l$ be a basis of $U \cap W$ and extend it to bases $u_1, ..., u_d$ of $U$ and $u_1, ..., u_l, w_{l + 1}, ..., w_d$ of $W$.
    As $\phi(U) = \phi(W)$, we can assume that the $u_i, w_i$ are scaled such that
    \begin{align*}
        0 =& (u_1 \wedge ... \wedge u_d) - (u_1 \wedge ... \wedge u_l \wedge w_{l + 1} \wedge ... \wedge w_d) \\
        =& u_1 \wedge ... \wedge u_l \wedge \bigl( (u_{l + 1} \wedge ... \wedge u_d) - (w_{l + 1} \wedge ... \wedge w_d) \bigr)
    \end{align*}
    By Lemma~\ref{prop:subspace_closed} we see that
    \begin{equation*}
        u_2 \wedge ... \wedge u_l \wedge \bigl( (u_{l + 1} \wedge ... \wedge u_d) - (w_{l + 1} \wedge ... \wedge w_d) \bigr) = 0
    \end{equation*}
    as $u_1 \notin \vspan\{u_2, ..., u_d, w_{l + 1}, ..., w_d\}$.
    Inductively, this argument shows that
    \begin{equation*}
        (u_{l + 1} \wedge ... \wedge u_d) - (w_{l + 1} \wedge ... \wedge w_d) = 0
    \end{equation*}
    If $l < d$, we can now apply Lemma~\ref{prop:subspace_closed} again to see that
    \begin{equation*}
        u_{l + 1} \in \vspan\{ u_{l + 2}, ..., u_d, w_{l + 1}, ..., w_d \}
    \end{equation*}
    as $u_{l + 2} \wedge ... \wedge u_d \neq 0$ by Lemma~\ref{prop:extpow_zero_iff_independent}.
    However, this contradicts the linear independence of $u_{l + 1}, ..., u_d, w_{l + 1}, ..., w_d$.
    Hence it must be $l = d$ and so $U = W$.
\end{proof}

\section{Part III}
In this part, we want to investigate the geometric properties of the Grassmanian resp. the image of $\phi$.
First of all, we introduce coordinates on $\Proj(\extpow^d k^n)$.
    Note that in the proof of Lemma~\ref{prop:isomorphism_extpow_det} it was shown that $v_{i_1} \wedge ... \wedge v_{i_d}$ for $i_1 < ... < i_d$ is a basis of $\extpow^d k^n$ if $v_1, ..., v_n$ is a basis of $V$.
    We introduce the homogeneous coordinates w.r.t. that basis, namely
        x: \Proj(\extpow^d k^n) &\to \Proj_k^{\{1, ..., n\}^{(d)}} \cong \Proj_k^{{n \choose d} - 1}, \\
        \Bigl[ \sum_{i_1 < ... < i_d} \lambda_{i_1, ..., i_d} (v_{i_1} \wedge ... \wedge v_{i_d}) \Bigr] &\mapsto \bigl[ \lambda_{i_1, ..., i_d} \bigr]_{i_1 < ... < i_d}
    The individual coordinates will be denoted by $x_I$ for some $I \in \{1, ..., n\}^{(d)}$ or $x_{i_1, ..., i_d}$ for $i_1 < ... < i_d$.
\begin{proposition}[3a]
    \label{prop:equations_grassmanian_2}
    For the embedding $\phi: \Gr(2, V) \to \extpow^2 V$ we have
        \Gr(2, V) \cong \mathrm{im}\phi = \V(I)
    where
        I := \langle x_{i, j}x_{u, v} + x_{i, v}x_{j, u} - x_{i, u}x_{j, v} \ | \ i < j < u < v \rangle \leq k\bigl[\Proj(\extpow^d V)\bigr] = k[x_{i, j} \ | \ i < j]
\end{proposition}
    By Lemma~\ref{prop:image_phi} we have that
        [\omega] \in \mathrm{im}\phi \ \Leftrightarrow \ \omega\ \text{decomposable}
    and so by Lemma~\ref{prop:characterization_decomposable_d_even}
        \omega \in \mathrm{im}\phi \ \Leftrightarrow \ \omega \wedge \omega = 0
    In $\Proj(\extpow^d V)$ we find that
        &\Bigl( \sum_{i < j} x_{i, j} (e_i \wedge e_j) \Bigr) \wedge \Bigl( \sum_{u < v} x_{u, v} (e_u \wedge e_v) \Bigr) = \sum_{\substack{i < j\\u < v}} x_{i, j} x_{u, v} (e_i \wedge e_j \wedge e_u \wedge e_v) \\
        =& \ 2\sum_{\substack{i < j < u < v}} x_{i, j} x_{u, v} (e_i \wedge e_j \wedge e_u \wedge e_v) + 2\sum_{\substack{i < u < j < v}} x_{i, j} x_{u, v} (e_i \wedge e_j \wedge e_u \wedge e_v) \\
        &+ 2\sum_{\substack{u < i < j < v}} x_{i, j} x_{u, v} (e_i \wedge e_j \wedge e_u \wedge e_v) \\
        =&2 \sum_{i < j < u < v} (x_{i, j} x_{u, v} - x_{i, u} x_{j, v} + x_{j, u} x_{i, v}) (e_i \wedge e_j \wedge e_u \wedge e_v)
    As the $e_i \wedge e_j \wedge e_u \wedge e_v$ are linearly independent, we see that for $[\omega] \in \Proj(\extpow^2 V)$ we have
        [\omega] \in \mathrm{im}\phi \ \Leftrightarrow \ \forall i < j < u < v: \ (x_{i, j} x_{u, v} + x_{i, v} x_{j, u} - x_{i, u} x_{j, v})(\omega) = 0
    Hence $\mathrm{im}\phi = \V(I)$.
\begin{example}[2b]
    For $n = 4$, Prop.~\ref{prop:equations_grassmanian_2} yields that $\Gr(2, 4) \cong \mathrm{im}\phi = \V(I)$ where
        I = \langle x_{1, 2} x_{3, 4} + x_{1, 4} x_{2, 3} - x_{1, 3} x_{2, 4} \rangle \in k[x_{1, 2}, x_{1, 3}, x_{1, 4}, x_{2, 3}, x_{2, 4}, x_{3, 4}]
    Changing the indices used for the coordinates, we find
        \Gr(2, 4) = \V(x_0 x_5 + x_2 x_3 - x_2 x_4)
    which is exactly what we found in the lecture.
\end{example}
\begin{example}[2c]
    For $n = 5$, Prop.~\ref{prop:equations_grassmanian_2} yields that $\Gr(2, 5) \cong \mathrm{im}\phi = \V(I)$ where
        I = \langle &x_{1, 2} x_{3, 4} + x_{1, 4} x_{2, 3} - x_{1, 3} x_{2, 4}, \quad x_{1, 2} x_{3, 5} + x_{1, 5} x_{2, 3} - x_{1, 3} x_{2, 5}, \\
        &x_{1, 2} x_{4, 5} + x_{1, 5} x_{2, 4} - x_{1, 4} x_{2, 5}, \quad x_{1, 3} x_{4, 5} + x_{1, 5} x_{3, 4} - x_{1, 4} x_{3, 5}, \\ 
        &x_{2, 3} x_{4, 5} + x_{2, 5} x_{3, 4} - x_{2, 4} x_{3, 5} \rangle \leq k\bigl[ \Proj(\extpow^2 k^5) \bigr]
    Using the following Sage-code, we can compute the number of intersection points of $\Gr(2, 5)$ with 3-dimensional hyperplanes, and find a probable value for its degree.
    \lstinputlisting{./examples.sage}
\end{example}
\end{document}
